---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "Pooja Savla"
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. Write code for a plot and describe what you observe?

```{r}
covid_data_count %>%
  filter(state == "MT") %>%
  ggplot(aes(x = date, y = Cases, color = race)) +
  geom_line(alpha = 0.7) +
  labs(
    x = "Date",
    y = "Cases",
    color = "Race",
    title = "COVID-19 Cases in Montana by Race"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.background = element_rect(fill = "lightgray"),  # Custom background color
    panel.grid.major = element_blank(),  # Remove major grid lines
    plot.title = element_text(hjust = 0.5)  # Center the plot title
  ) 

```

2. Are there any conclusions you can draw?

Here are the conclusions drawn from the plot:

Firstly, In January, there was a notable spike in COVID-19 cases within the "Black" racial community.
Secondly, The racial group with the highest number of COVID-19 cases overall was the "White" community.
Thirdly,  A significant surge in COVID-19 cases occurred within the "White" community during the period spanning November to December.
Fourthly, Between April and July, there were no substantial increases in COVID-19 cases, with case numbers remaining relatively low. A linear increase in cases was observed from July to September, followed by a departure from linear growth after October, indicating a more varied trend in total COVID-19 cases during that time.


3. What were you not able to do due to not having the R knowledge?

As a beginner in R, I successfully filtered the data for the state of Montana in the provided code. However, my proficiency in advanced data manipulation techniques, such as reshaping data, joining datasets, or handling missing values, is limited. Additionally, while I added a smoothed line with confidence intervals to the plot, my understanding of the statistical methods behind it and my knowledge of performing more advanced statistical analyses in R are currently limited.

4. What other data would be useful to better understand this data?

To gain a more comprehensive understanding of the COVID-19 situation, it would be beneficial to gather additional types of data:

1. **Demographic Information**: 
Collecting data on the population's demographics, such as 
age, 
gender, 
socioeconomic status, and 
underlying health conditions, 
can help pinpoint vulnerable groups and shed light on how COVID-19 affects various segments of society.

2. **Vaccination Records**: 
Compiling data related to COVID-19 vaccinations, including:
the total number of vaccinated individuals, 
vaccination rates across different demographic groups, and 
details about the types of vaccines administered, 
can provide valuable insights into the success of vaccination efforts.

3. **Testing Records**: 
Gathering information about COVID-19 testing, including:
the total number of tests conducted, 
the methods used for testing, 
and the outcomes (positive, negative, or inconclusive), 
is crucial for assessing the capacity and accuracy of the testing infrastructure.

